---
title: "Data Cleaning Example for VecDyn: Manatee County Monitoring Data Cleanup in R"
author: "Matthew J Watts"
date: "21 May 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Load the following packages (and install if required)

```{r}
#install.packages("tidyverse")
#install.packages("readr")
#install.packages("plyr")
#install.packages("scales")
library(tidyverse)
library(readr)
library(plyr)
library(scales)

```


###Load up the VecDyn template / dataframe and name it 'vecdyn_manatee_county_mosquito'.

```{r}
vecdyn_manatee_county_mosquito  <-data.frame(title = character(),
                             taxon = character(),
                             sample_start_date = character(),
                             sample_start_time = character(),
                             sample_end_date =  character(),
                             sample_end_time = character(),
                             value = numeric(),
                             sample_location_info = character(),
                             sample_lat_DD = character(),
                             sample_long_DD = character(),
                             sample_info = character(),
                             sample_name = character(),
                             sample_sex = character(),
                             location_description = character(),
                             location_environment = character(),
                             study_lat_DD = character(),
                             study_long_DD = character(),
                             spatial_accuracy = character(),
                             location_extent = character(),
                             geo_datum = character(),
                             species_id_method = character(),
                             study_design = character(),
                             sampling_strategy = character(),
                             sampling_method = character(),
                             sampling_protocol = character(),
                             measurement_unit = character(),
                             study_collection_area = character(),
                             value_transform = character(),
                             stringsAsFactors=FALSE)
#write.csv(VecDyn_template, file = "VecDyn_template.csv", row.names = FALSE)
                             
```

###Import the Manatee County Mosquito Data Sets (2012-2016) from the following URLs, when importing datasets, make sure you set 'stringsAsFactors=FALSE' to ensure R doesnt apply any unwanted formatting to the dataframe. 

```{r}
Manatee_County_Mosquito_2012 <- read.csv(url("https://zenodo.org/record/1217702/files/VectorBase-2012-Manatee_County_Mosquito_Control_District_Florida_USA.csv?download=1"), stringsAsFactors=FALSE)

Manatee_County_Mosquito_2013 <- read.csv(url("https://zenodo.org/record/1217704/files/VectorBase-2013-Manatee_County_Mosquito_Control_District_Florida_USA.csv"), stringsAsFactors=FALSE)

Manatee_County_Mosquito_2014 <- read.csv(url("https://zenodo.org/record/1217706/files/VectorBase-2014-Manatee_County_Mosquito_Control_District_Florida_USA.csv?download=1"), stringsAsFactors=FALSE)

Manatee_County_Mosquito_2015 <- read.csv(url("https://zenodo.org/record/1217709/files/VectorBase-2015-Manatee_County_Mosquito_Control_District_Florida_USA.csv?download=1"), stringsAsFactors=FALSE)

Manatee_County_Mosquito_2016 <- read.csv(url("https://zenodo.org/record/1217711/files/VectorBase-2016-Manatee_County_Mosquito_Control_District_Florida_USA.csv?download=1"), stringsAsFactors=FALSE)
```

###Bind all the datasets together using  the rbind function
```{r}

Manatee_County_Mosquito <- rbind(Manatee_County_Mosquito_2012,Manatee_County_Mosquito_2013,Manatee_County_Mosquito_2014,Manatee_County_Mosquito_2015,Manatee_County_Mosquito_2016)


```

###Check the dataframe to see if everything went to plan

```{r}
head(Manatee_County_Mosquito)
```

###You can now remove all the unused data frames 
```{r}

rm(Manatee_County_Mosquito_2012,Manatee_County_Mosquito_2013,Manatee_County_Mosquito_2014,Manatee_County_Mosquito_2015,Manatee_County_Mosquito_2016)


```


###Next we need to extract values from each column and name each list according to its matching column name in vecdyn manatee_county mosquito data set (vecdyn template)

```{r}

taxon <- Manatee_County_Mosquito$species

value <- Manatee_County_Mosquito$sample_count

sample_sex <- Manatee_County_Mosquito$sex

sample_info <- Manatee_County_Mosquito$developmental_stage

sample_start_date <- Manatee_County_Mosquito$collection_date_start

sample_end_date <- Manatee_County_Mosquito$collection_date_end

sample_lat_DD <- Manatee_County_Mosquito$GPS_lat

sample_long_DD <- Manatee_County_Mosquito$GPS_lon

sample_location_info <- Manatee_County_Mosquito$location

location_description <- Manatee_County_Mosquito$location_ADM2

sampling_method <- Manatee_County_Mosquito$attractant

sampling_protocol <- Manatee_County_Mosquito$trap_type

species_id_method <- Manatee_County_Mosquito$identification_method

```

###Next we bind all the extracted values (lists) together and add them to the vecdyn_manatee_county_mosquito dataframe

```{r}

a <- cbind(taxon, value, sample_sex, sample_info, sample_start_date, sample_end_date, sample_lat_DD, sample_long_DD, sample_location_info, location_description, sampling_method, sampling_protocol, species_id_method)

a <- data.frame(a, stringsAsFactors=FALSE)

b <- rbind.fill(vecdyn_manatee_county_mosquito,a)
```

###Check and convert all dates to the Y-m-d format, in the case of this data set the dates are already in the correct format

```{r}
vecdyn_manatee_county_mosquito <- 
  b %>%
  dplyr::mutate(sample_end_date = as.Date(as.character(sample_end_date, format = "%Y/%m/%d"))) %>%
  dplyr::mutate(sample_start_date = as.Date(as.character(sample_start_date, format = "%Y/%m/%d"))) %>%
  dplyr::mutate(value = as.numeric(value))


```

###We next need to create specific data sets for VecDyn, in this case we will create a data set for species ' aedes aegypti'. We will also give the data set a title and provide details on the measurement unit

```{r}

vecdyn_manatee_county_mosquito_Aedes_aegypti <- 
  vecdyn_manatee_county_mosquito %>%
  dplyr::filter(taxon == "Aedes aegypti")  %>%
  dplyr::mutate(title = 'Manatee County Monitoring Aedes aegypti') %>%
  dplyr::mutate(measurement_unit = 'individuals') ##it is always good to keep track of the measurement unit


```

### Test the data set by producing a time series plot in ggplot. Note that this particular data set contains lots of sampling stations within manatee_county

```{r}

vecdyn_manatee_county_mosquito_Aedes_aegypti %>%
  dplyr::select(value, sample_end_date) %>%
  dplyr::group_by(sample_end_date) %>%
  dplyr::summarise(value = sum(value)) %>%
  ggplot(aes(sample_end_date, value)) + geom_line() +
  scale_x_date(labels = date_format("%b/%y"))
  

```



###Finally write the data set to CSV, set all missing value to blank ""

```{r}
write_csv(vecdyn_manatee_county_mosquito_Aedes_aegypti,"vecdyn_manatee_county_aedes_aegypti.csv", na = "")
```

###Lets also make a data set for Anopheles quadrimaculatus

###Again We need to create specific data sets for VecDyn, in this case by species 'Anopheles quadrimaculatus'

```{r}

vecdyn_manatee_county_mosquito_A.quadrimaculatus <- 
  vecdyn_manatee_county_mosquito %>%
  dplyr::filter(taxon == "Anopheles quadrimaculatus")  %>%
  dplyr::mutate(title = 'Manatee County Monitoring Anopheles quadrimaculatus') %>%
  dplyr::mutate(measurement_unit = 'individuals') ##it is always good to keep track of the measurement unit

View(vecdyn_manatee_county_mosquito_A.quadrimaculatus)
```

### Test the data set by producing a time series plot in ggplot. Note that this particular data set contains lots of sampling stations within manatee_county

```{r}

vecdyn_manatee_county_mosquito_A.quadrimaculatus %>%
  dplyr::select(value, sample_end_date) %>%
  dplyr::group_by(sample_end_date) %>%
  dplyr::summarise(value = sum(value)) %>%
  ggplot(aes(sample_end_date, value)) + geom_line() +
  scale_x_date(labels = date_format("%b/%y"))
  

```


###Finally write the data set to CSV, set all missing value to blank

```{r}
write_csv(vecdyn_manatee_county_mosquito_A.quadrimaculatus,"vecdyn_manatee_county_a.quadrimaculatus.csv", na = "")
```
